### 2012-07-10 4.tex
Writing on chapters 2, 4

%\chapter{Structural and accidental gaps\protect\symbolfootnote[1]{An abridged version of this chapter will appear in the proceedings of the 30th West Coast Conference on Formal Linguistics under the same name.}}
\label{clusters}

The term \emph{lexicon} has many senses: throughout, I use it to refer simply to the set of underlying representations.
%\citep[][269]{LANGUAGE}.

In the preceding chapter, it was proposed that

Reconsider the null hypothesis presented in the preceding chapter:

\begin{example}[null hypothesis]
Universal grammar does not countenance constraints on the contents of URs
\end{example}

The idea

relationship to proposal

One primary way which this null hypothesis has been critiqued is by demonstrating that the lexicon of some language 
exhibits dispreferences or gaps corresponding to.

argue by trends in the lexicon

possibility of sparsity (C\&H, Fischer/Vogt and recent stats)

Of course, to obtain the desired results, we must guarantee that each sequence structure rule reflects a general systematic fact about the language, and not a fact which is due merely to the existence of accidental gaps in the lexicon. \citep[][401, fn.~8]{Stanley1967}

In the preceding chapter, it was proposed that

Lexicon

This appears to be exceptionlessly true of Arabic

That even gradient patterns

One of the first study of this type was an analysis of co-occurrence restrictions on consonants in Javanese roots by \citet{Mester1988}, recently revisited by \citet{Graff2011}. 
Another Austronesian language studied in this fashion is Muna \citep{Coetzee2008a,Anttila2008}.
This technique has been applied to consonants in Semitic, especially Arabic \citep{McCarthy1988,McCarthy1994,Pierrehumbert1993,Frisch1996,Frisch2004,Coetzee2008a} but also in Tigrinya \citep{Buckley1997}, Hebrew \citep{Berent2003}, and Amharic \citep{Colavin2010}, and to various co-occurrence restrictions in English \citep{Berkley1994b,Berkley1994a,Pierrehumbert1994,Dmitrieva2008a,Dmitrieva2008b,Coetzee2008b}, Russian \citep{Padgett1992}, Dutch \citep{Graff2011}, Navajo \citep{Martin2007,Martin2011} and Gitksan \citep{Brown2010}.

The absence of 
The claim here is that however stark the absence or underrepresentation of some form is, it

This hypothesis admits the possibility that there may be accidental gaps in the lexicon, a possibility which predates generative thinking:

\begin{quote}
\ldots{}the fact that some [clusters--KG] are not found must be due to accidental gaps in the inventory of signs, and cannot be explained by structural laws of the language. \citep[][16]{Fischer-Jorgensen1952}
\end{quote}

\noindent
Hans \citeauthor{Vogt1954} makes a similar observation in his study of Georgian clusters

\begin{quote}
Although my material is drawn from a fairly extensive corpus---all accessible dictionaries and vocabularies, printed texts of tens of thousands of pages as well as ordinary speech---there is every reason to believe, as experience has shown, that additional material would yield new clusters. The material will never be complete. It will always contain accidental gaps \ldots partly because some clusters by pure chance do not occur in the vocabulary. \citep[][30]{Vogt1954}
\end{quote}

\citet{Chomsky1965}

This chaper demonstrates that apparently systematic gaps may arise even in the absence of phonotactic preferences, and that this possibility compromises attempts to show that such preferences shape the lexicon.

\section{Domain}

% WHY I'm DOING THIS STUDY
% describes makeup of the corpus and how p

\subsection{Data sources}

The English CELEX database \citep{CELEX},

\citet{Aronoff1976}

\citet{Harley2009}

\citet{Taft1975} ?
\citet{Taft1981} ?
\citet{Taft2004a} ?

\citet{Halle1962}

% 2.1.4: k

Under the parses and phonologization described immediately below, the 6,876 simplex CELEX entries contain 21 medial coda types and 40 unique medial onsets. Of the 840 ($= 21 \times 40$) possible syllable contact clusters that could be produced by free combinations of the attested medials codas and onsets, 158 are attested, a saturation rate of 18.8\%.

\subsection{Syllabification}

For this study, it is necessary to separate medial consonant clusters into codas and onsets. While CELEX provides syllabified transcriptions, no description of the syllabification procedure is given in the documentation, and inspection of these syllabifications suggests the procedure used is not fully systematic. For instance, \emph{chemistry} is syllabified as [ˈkɛ.mɪ.strɪ] but \emph{ministry} as [ˈmɪ.nɪs.trɪ].\footnote{Note that word-final \emph{y} is generally lax [ɪ] in Received Pronunciation \citep[][294]{AOE2}.} This putative [ɪ.strɪ $\sim$ ɪs.trɪ] contrast, and many other such contrasts in the CELEX data, are dubious simply because there is no clear evidence that contrastive syllabification occurs in any language. Apparent counterexamples appear to be effects of vowel or consonant length \citep[e.g.,][]{Elfner2006}, word stress, or morphological structure, none of which explain the \emph{chemistry}/\emph{ministry} syllabification contrast.

An automated syllabification system was constructed and applied to medial clusters in the CELEX data. This system need only deal with simplex English words that contain the medial consonant clusters which are the focus of this study.  There is, for instance, no need to address the status of ``ambisyllabic'' consonants, i.e., singleton medial consonants preceded by a stressed lax vowel \citep[][219f.]{Rubach1996}, or to address morphological effects on syllabification. The technique used is a variation on the theme of onset maximization (e.g., \citealt{Kurylowicz1948}, 
\citealt[75]{Pulgram1970}, \citealt[42f.]{Kahn1976}, \citealt[][358f.]{Selkirk1982b}) which parses word-medial clusters so that the onset is as large as possible. Medial clusters in words like \emph{neu}[.tr]\emph{on} or \emph{bi}[.str]\emph{o} are found in word-initial position (e.g., [tr]\emph{ansit}, [str]\emph{ike}), and onset maximization leaves the coda of the first syllable empty. In contrast, the [nstr] cluster in \emph{mi}[n.str]\emph{el} does not occur word-initially; here the maximal onset is [str], and the [n] is assigned to the coda.

There is one case where unchecked maximization of medial onsets produces incorrect syllabifications. When a medial consonant cluster is preceded by a stressed lax vowel, as in words like \emph{propaga}[n.d]\emph{a}, \emph{whi}[s.p]\emph{er}, \emph{vi}[s.t]\emph{a}, or \emph{bi}[s.k]\emph{uit}, the first consonant of the cluster checks the lax vowel \citep[e.g.,][3]{Hammond1997}. \citet[][55]{Harris1994} notes that onset maximization makes the wrong predictions when the medial cluster is a valid onset: in \emph{whisper}, \emph{vista}, and \emph{biscuit}, it incorrectly assigns [sp, st, sk] to the medial onset, leaving the coda empty. The approach adopted here is essentially that of \citet{Pulgram1970}: the first consonant of a medial cluster is assigned to the coda of a preceding coda if is immediately preceded by a stressed lax vowel, which do not occur syllable-finally in English.\footnote{\citet[][589f.]{Lowenstamm1981} observes a strikingly similar issue with the onset maximization principle in French, and proposes a similar solution.}

If the English affricates [tʃ, dʒ] are treated as sequences and not individual segments, this addition to the standard onset maximization technique would itself make the wrong prediction for medial affricates preceded by lax stressed vowels in words like \emph{ra}[.tʃ]\emph{et} or \emph{a}[.dʒ]\emph{ile}, incorrectly assigning the stop and fricative portions to separate syllables. For this reason, an additional constraint is assumed which prevents the two components of the affricates from being split by syllabification. This is motivated by the tendency of affricates to pattern with single segments in many languages. For instance, the only complex onsets in Classical Nahua are the affricates [ts, tʃ, tɬ] \citep[][9]{Launey2011}.

\subsection{Inventory consdierations and subsyllabic parsing}

To identify the inventory of syllable contact clusters, it is also necessary to parse syllables into onset, nucleus, and coda. In most cases this is trivial, but in a few cases, it is necessary to use distributional heuristics to determine the underlying inventory and subsyllabic affiliations.

\subsubsection{The velar nasal} \label{velarnasal}

There is a long-standing debate regarding whether English [ŋ] is a phoneme in its own right \citep[e.g.,][]{Sapir1925}, an analysis which is a consequence of principles like the Alternation Condition \citep{Kiparsky1968} or Lexicon Optimization \citep[][53]{OT}, or simply the allophone of /n/ found before /k, g/ (e.g., \emph{SPE}:85, \citealt{Borowsky1986}:65f.). The strongest piece of evidence for the latter position, in which the velar nasal is a pure allophone, is the general absence of onset [ŋ], a position where it can never be followed by a dorsal consonant needed to derive the velar allophone. \citet{Pierrehumbert1994} assumes the pure allophonic analysis in her study of syllable contact clusters, and it is assumed here. English nasal place allophony is formalized in Section \ref{cnpasection} below.

\subsubsection{[j] onglides}

The [j] onglide in words such as in words such as \emph{ass}[j]\emph{ume} is assumed to be present in underlying representations (e.g., \citealt[][278]{Borowsky1986}, J. \citealt{Anderson1988b}, pace \emph{SPE}:196, \citealt[][89]{Halle1985a}, \citealt[][217]{McMahon1990}), since the presence or absence of the glide is at least marginally contrastive (e.g., \emph{coo}/\emph{coup} \alt{} \emph{queue}, \emph{booty} \alt{} \emph{beauty}). The front onglide is further assumed to be assigned to the nucleus, except when the onset would otherwise be null (e.g., \emph{jun}[j]\emph{or}).

There is considerable evidence for this assumption. When [j] is a simplex onset, it may be followed by any vowel \citep[][276]{Borowsky1986}, but when [j] is immediately preceded by an onset consonant (e.g., [bj]\emph{ugle}), the following vowel is always [uː]. In fact, speakers judge words in which the front onglide is preceded by an onset consonant and followed by a vowel other than [uː] to be anomalous \citep{Moreland2009}. This defective distribution of vowels following [j] suggests that the onglide in this context is the first component of a phonological diphthong (\citealp[][232]{Hayes1980}, \citealp[][61f.]{Harris1994}, \citealp{Davis1995}). \citet[][42]{Clements1983} note that /m, v/ do not appear in onset clusters except in words like \emph{muse} or \emph{view}; the exceptionality of [juː] also suggests the glide is nuclear. There is also strong external support for this principle. The [juː] in words like \emph{spew} may pattern together in Pig Latin \citep{Davis1995,Idsardi2005} and \emph{shm}-reduplication \citep{Nevins2003} to the exclusion of the rest of preceding tautosyllabic consonants. The same fusion of [juː] at the expense of the onset is also found in speech errors, e.g., [kjuː]\emph{mor} [h]\emph{omponent} for [hjuː]\emph{mor component} \citep[][130]{Shattuck-Hufnagel1986}.

\subsubsection{[w] onglides}

The selective properties of the back onglide [w] contrast sharply with those of the front onglide, and I assume that it is assigned to the onset. Whereas the front onglide shows only limited selectivity for preceding tautosyllabic consonants \citep{Davis1995,Kaye1996}, the back onglide [w] is rarely preceded by tautosyllabic consonants other than [k] (e.g., \emph{tran}[kw]\emph{il}). Unlike the front glide, syllable-initial [kw] may be followed by nearly any vowel \citep[][161]{Davis1995}. Unlike [juː], onglide [w] followed by a vowel does not pattern together in Pig Latin \citep[][166]{Davis1995}. And onsets followed by [w] pattern with other complex onsets in undergoing epenthesis in the patholical speech of VBR, discussed immediately above.

\subsubsection{Post-vocalic \emph{r}}

Both CELEX and the dictionary used by \citet{Pierrehumbert1994} in her study transcribe Received Pronunciation, in which word-medial post-vocalic \emph{r} has been lost. In \emph{r}-full dialects, there is reason to believe that post-vocalic \emph{r} is in fact nuclear, and therefore its presence or absence is irrelevant to the constraints on syllable contact clusters. Many vowel contrasts are suspended before \emph{r} (e.g., \citealt[269f.]{Fudge1969}, \citealt[][255]{Harris1994}); for example, \emph{fern}, \emph{fir}, and \emph{fur} lack the contrasts found in \emph{pet}, \emph{pit}, and \emph{putt}. Further evidence for the nuclear status of post-vocalic \emph{r} comes from variable phonological processes in which post-vocalic \emph{r} patterns differently than internal coda consonants. \citeauthor{Harris1994} reports that a variable process of /t/-\textsc{Glottalization} in many dialects of British English is blocked when /t/ is preceded by any consonant except post-vocalic \emph{r}.

\begin{example}[/t/-\textsc{Glottalization} in British English (after \citealp{Harris1994}:195, 258)]
\begin{tabular}{l l l@{} l}
a. & par[t]   &   & par[ʔ]   \\
   & car[t]on &   & car[ʔ]on \\
b. & fis[t]   & * & fis[ʔ]   \\
   & mis[t]er & * & mis[ʔ]er \\
\end{tabular}
\end{example}

\noindent Similarly, while /t, d/ delete in word-final position when immediately preceded by a consonant, including sonorants /n, l/, as in (\ref{td}a), deletion of /t, d/ after post-vocalic \emph{r} in American English is ``rare or nonexistent'' \citep[][8]{Guy1980}.

\begin{example}[/t, d/-\textsc{Deletion} in American English] \label{td}
\begin{tabular}{l l l@{} l}
a. & be[lt]  &   & be[l] \\
   & me[nd]  &   & me[n] \\
b. & sh[ɚt]  & * & sh[ɚ] \\
   & c[ɚd]   & * & c[ɚ]  \\
\end{tabular}
\end{example}

@@

\subsection{Static and derived constraints}

With this database of English syllable contact clusters it is now possible to evaluate constraints on sequences of underlying phonemes.

Here, the metric used is simply

The evaluation

The foremost evidence for

The evaluation used here is simply

Having described the database of English syllable contact clusters,

I now consider constraints on sequences of URs as predictors of attested and unattested contact clusters. First, I describe the statistical technique used, then apply this technique to both phonologically derived constraints and a set of ``static'' phonotactic constraints proposed by \citet{Pierrehumbert1994}.

Before the dawn of generative phonology, many linguists concerned themselves with documenting co-occurrence in various languages. \citet[][28]{Vogt1954} declares that a ``phonemic description of a language \ldots should also comprise a description of the phoneme combinations that occur or may be expected to occur in the language''. Studies in this tradition include the discussions of consononantal co-occurrence in Arabic and Javanese by \citet{Greenberg1950} and \citet{Uhlenbeck1950}, respectively. \citeauthor{Uhlenbeck1950}, for instance, puts forth co-occurrence dispreferences, rather than exceptionless generalizations of the sort given by \citet{Greenberg1950}.

\citet{McCarthy1988}

Years later, \citet{Mester1988} revisits \citeauthor{Uhlenbeck1950}'s study and proposes a statistical technique, based on the chi-square test, to distinguish between accidental and structural lexical dispreferences in the Javanese lexicon.

\citeauthor{Mester1988}'s technique has been adopted by many subsequent studies (e.g., \citealt{Padgett1992,Padgett1995} on Russian; \citealt{Pierrehumbert1993}, \citealt{Frisch1996}, and \citealt{Frisch2004} on Arabic; \citealt{Berkley1994b,Berkley1994a,Berkley2000}, \citealt{Dmitrieva2008a}, \citealt{Dmitrieva2008b} on English). The null hypothesis $H_0$ is that some sequence occurs at no different a rate than the researcher expects, and the alternative hypothesis $H_1$ is that it is underrepresented. Given an observed frequency $O$ and also a frequency $E$ expected under the null hypothesis, the test is as follows.

\begin{example}[null hypothesis testing]
$\displaystyle \textrm{Reject } H_0 \textrm{ iff } \frac{(O E) ^ 2}{E} >
χ^2$
\end{example}

%\footnote{This is the one-tailed version of the test.}

In this study, I use the \citet{Fisher1934} exact test, which is isomorphic to this chi-square test. Compared to the chi-square test, the Fisher exact test $p$-value is somewhat more difficult to compute, but both are automated by modern statistical packages and can be computed rapidly. The major advantage of the Fisher test over the chi-square test is that it is appropriate for small samples or rare events, whereas the chi-square test depends on an approximation which is only exact in the limit (i.e., with a hypothetically infinite sample), and therefore is not appropriate for small samples.

On the other end of the spectrum, the Monte Carlo statistical procedure developed by \citet{Kessler2001} and applied to co-occurrence statistics by \citet{Martin2007,Martin2011} and \citet{Brown2010} does not scale to large samples. The Monte Carlo procedure is isomorphic to the one-tailed Fisher Exact test, and involves comparing observed co-occurrence counts to those produced by randomly generated samples.  This requires the researcher to generate random permutations, something which is particularly difficult for large samples, as I will explain; disinterested readers may wish to skip the following paragraph.

Any permutation of a sequence $L$ can be represented as a sequence $S$ of the same length, where the value of $S_i$ corresponds to the permuted position of $L_i$. Generating random permutations requires an unbiased method to generate all $N!$ of these lists with equal probability. Unfortunately, this is near impossible for $N$s of moderate size with current computational resources. Any pseudo-random number generator can be characterized by its ``period'', the number of pseduo-random numbers it can emit before repeating itself; this is also the upper limit for the number of random permutations that can be generated. The pseudo-random number generator used by both \citeauthor{Martin2011} and \citeauthor{Brown2010} has a period of approximately $2^{48}$, a number which turns out to be smaller than $17!$. For the large samples, sometimes in the thousands, used by these researchers, virtually all the possible permutations can never be generated by this method. As a consequence, the shuffling procedure is far from random and introducing the possibility that the statistical test is biased in unpredictable ways.

As an example of the Fisher exact test as it is applied here, consider some of the Javanese co-occurrence facts considered by \citet{Mester1988}. \citeauthor{Mester1988} is at pains to show that there is a gradient dispreference for first and second root consonants which share the same major place. However, perfectly identical first and second consonants are exempt. Since

% APPLY THIS TO JAVANESE

\begin{example}[Javanese root co-occurrence] %(\citealp[][264]{Uhlenbeck1950}, \citealp[][139]{Mester1988})]
\begin{tabular}{l r r r r}
\toprule
           & attested & unattested & saturation & $p$-value \\
\midrule
conforming & 134      & 12         & 0.918      & \multirow{2}{*}{6.095\e{-06}} \\
violating  & 21       & 15         & 0.583 \\
\bottomrule
\end{tabular}
\end{example}

\noindent
It is evident that roots with identical labial stops are far more common than those in which the roots disagree in voicing.

This data is input to a chi-square test, which reports that this highly skewed distribution is unlikely to be due to chance

That this is highly unlikely to be due to chance is apparent both from the chi-square ($\chi^2 = 25.572$, two-tailed $p = 4.3$\e{-07}) and Fisher exaact tests ($p =  6.095$\e{-06}).

I use two-tailed statistical tests throughout. 
Under a one-tailed test, the alternative hypothesis is underrepresentation, whereas a two-tailed test also allows the possibility that a class of clusters would be overrepresented, i.e., more frequent than predicted by the null hypothesis. As

\citet{Pierrehumbert1994}

It must be stressed that under current assumptions, this is independent of whether predictable prosodic representations (particularly coda and onset identity) are part of the memorized underlying representation \citep[e.g.,][]{Vaux2003}.
It is an undeniable fact that URs need to be 
The derivative nature of 
\citet{Ito1989a,Noske1992}

least some researchers (i.e., \citealt{Mester1988})

 \citep[e.g.,]{Brown2010} have argued for phonotactic constraints giving rise to overrepresentation, so I admit this possibility here and use two-tailed tests throughout.

%\footnote{For example, \citet{Martin2007} applies this technique to a list of 4,758 noun-noun compounds. The number of permutations of a list of this length can be estimated using Stirling's approximation.

%\ex $ \lim_{n \rightarrow \infty} n! = e ^ {n \ln{n} - n} = 2 ^ {\log_2{e} (n \ln{n} - n)} $ \xe

%\noindent
%$4,758!$ is approximately $2 ^ {51260}$. The author is unaware of any system that provides 51,260 bits of entropy; and some popular programming languages, such as ANSI C, provide as few as 15 bits.}

%%% START OF SERIOUSNESS

\subsection{Derived constraints}

\emph{SPE} describes three phonological alternations which target syllable contact clusters. The corresponding constraints on simplex syllable contact clusters are considered below.

\subsubsection{Obstruent voice assimilation}

Voice assimilation in English is exemplified by the allomorphs of /-d/ suffix, forming regular pasts and denominal adjectives, and the /-z/ suffix, forming regular noun plurals, genitives, and the 3sg. present active indicative.\footnote{The URs /-d, -z/ are argued for by \citet[][282]{Hockett1958}, \citet[][210]{SPE}, \citet{Basboll1972}, \citet{Shibatani1972}, S. \citet[][]{Anderson1973a}, \citet[][102]{Pinker1988}, and \citet[][284f.]{Bakovic2005b}; alternative URs are argued for by \citet[][210f.]{LANGUAGE}, \citet[][426]{Nida1948}, \citet{Luelsdorff1969}, \citet{Lightner1970}, \citet{Hoard1971}, \citet[]{Miner1975}, \citet{Zwicky1975}, \citet{Kiparsky1985}, and \citet[][135]{Borowsky1986}.}

\begin{example}[Inflectional affix voice assimilation]
\begin{tabular}{l l l l l l}
a. & nap & nap[t] \\
   & nab & nab[d] \\
b. & lap & lap[s] \\
   & lab & lab[z] \\
\end{tabular}
\end{example}

\noindent The devoicing of the two voiced obstruent suffixes can be formalized as a rule spreading the \textsc{Voice} specification of an obstruent rightward.

\begin{example}[\textsc{Obstruent Voice Assimilation} (after \emph{SPE}:178)]
\xymatrix@R=24pt@C=24pt{
\txt{[α \textsc{Voice}]}\ar@{-}[d]\ar@{--}[dr] &                               \\
\txt{C}\ar@{-}[d]                              & \txt{C}\ar@{-}[d]             \\
\txt{[$+$\textsc{Obstruent}]}                  & \txt{[$+$\textsc{Obstruent}]} \\
}
\end{example}

\citet{Pierrehumbert1994} does not mention voice assimilation in her study of English syllable contact clusters. \citet[][74f.]{Hammond1999a} notes the existence of words like \emph{a}[b.s]\emph{inth} and \emph{a}[s.b]\emph{estos}, which contain obstruent voice clusters disagreeing in [\textsc{Voice}]. This might be taken as evidence that \textsc{Obstruent Voice Assimilation} does not apply to simplex words, and indeed, the \textsc{Revised Alternation Condition} (RAC) proposed by \citeauthor{Kiparsky1973a} (\citeyear{Kiparsky1973a}:163, \citeyear{Kiparsky1982a}:152) blocks the application of obligatory neutralization processes like \textsc{Obstruent Voice Assimilation} in root-internal (i.e., non-derived) environments.\footnote{The ``obligatory'' condition on the RAC has the result that \textsc{Obstruent Voice Assimilation} is always consistent with the RAC whether or not it applies in non-derived environments. If the rule applies in non-derived environments, then it has lexical exceptions (e.g., \emph{a}[s.b]\emph{estos}) and is not obligatory and thus free from the RAC. On the other hand, if the rule is subject to the RAC, it only applies in derived environments, in which it is obligatory.} Despite this, root-internal clusters of obstruents which disagree in voicing (e.g., [s.b, b.s]) are much rarer than those which contain adjacent voiced or voiceless obstruents (e.g., [s.p, b.z]), suggesting that \textsc{Obstruent Voice Assimilation} applies in non-derived environments, preventing learners from positing URs in which adjacent obstruents do not agree in voice.

>>>>

\chapter{Structural and accidental gaps} \label{clusters}

Early work in generative phonology recognized that neutralizing, surface-true phonological alternations impose constraints on surface forms and underlying representations (\citealp[283]{Anderson1974}, \citealp[382]{SPE}, \citealp[205f.]{Dell1973}, \citealp[22f.]{SPR}, \citealp{Kenstowicz1977}: chapter 3, \citealp[28f.]{Stampe1979}, \citealp[410f.]{Stanley1967}). Despite isolated claims to the contrary (e.g., \citealp[297]{Hale1965}; \citealp[212f.]{Postal1968}), however, the \emph{communis opinio} 
(including many of the authors cited above) holds that not all phonotactics---that is, language-specific sound sequence co-occurrence restrictions---can be derived in this fashion, and phonotactic knowledge is independent of the phonology.

\begin{quote}
It would be a result of the greatest interest if it were to turn out that every intramorphemic regularity was necessarily reflected in an alternation, and thus in a rule; but this position cannot seriously be maintained. \citep[283]{Anderson1974}
\end{quote}

English word-medial consonant clusters have been used, by \citet{Pierrehumbert1994} and others, to argue for the necessity of an independent phonotactics module and illustrate proposals for its architecture. This chapter argues, however, that the only structural constraints on the concatenation of codas and onsets to form word-medial clusters are derivative of well-documented phonological alternations in English.

\section{Background}

English admits a variety of word-medial consonant clusters spanning syllable boundaries, known as \emph{syllable contact clusters} or \emph{interludes}. These clusters can be as short as two consonants (e.g., \emph{a}[n.t]\emph{ics}; singleton intervocalic consonants are ignored here) or as long as four (e.g., \emph{mi}[n.str]\emph{el}). \citeauthor{Pierrehumbert1994} presents the following as a null hypothesis for which clusters are admissible in any given language \citep[cf.][]{Haugen1956}.

\begin{quote}
\ldots{}in the absence of additional provisos, any concatenation of a well-formed coda and a well-formed onset is predicted to be possible medially in a word. \citep[][168]{Pierrehumbert1994}
\end{quote}

This chapter is concerned with the nature of these ``provisos'' as reflected by which of the ``possible'' clusters, so defined, are attested. While many phonotactic generalizations can be thought of as inventory restrictions imposed by prosodic constituents, additional filters on the set of possible clusters are essentially combinatoric in nature and thus distinct, there is no prosodic constituent consisting of the coda of one consonant-final syllable and the onset of a following consonant-initial syllable \citep[though see][]{Steriade1999}.

@@

\citeauthor{Pierrehumbert1994} argues that the inventory of clusters in English shows the effect of a small number of static phonotactic constraints. The remainder of this chapter provides an exhaustive quantitative evaluation of this hypothesis. The most important constraints on medial clusters are those derived from phonological rules which apply across syllable boundaries; there is no remaining role for static constraints. The sparse nature of the lexicon plays a secondary role, giving rise to apparent phonotactic gaps even in the absence of static phonotactic constraints.

\subsection{Corpus construction}

A list of medial clusters is gathered by applying a number of filters and transformations to an English lexical database. The resulting list of clusters is reproduced in Appendix \ref{appendixB} and in electronic form at the author's website.

\subsubsection{Simplex words}

An analysis of lexical constraints requires a corpus of lexical representations. So as to remain agnostic about particulars of morphological theory, the goal here is to identify words which bear no overt inflection, and which do not admit any plausible further decomposition, henceforth ``simplex words''. For the structuralists, a procedure for parsing words into morphs, just the thing for locating simplex words, held the status of ``philosopher's stone'', a sort of near-mythical objective of the endeavor, but actual attempts to develop such a procedure \citep[e.g.,][]{Harris1955,Nida1948} are now recognized as anything but theory-neutral. Yet, there is no alternative to adopting some computational procedure or human-coded database as the ``gold standard'' for a study of this type. \citet{Pierrehumbert1994} analyzes a list of entries in the Collins English dictionary by hand, using her intuitions to filter out ``complex'' words as well as those which are not ``reasonably familiar''. Unfortunately, the resulting list was neither published nor circulated, and the possibility of replicating \citeauthor{Pierrehumbert1994}'s' sensations of morphological complexity is remote. Following prior studies of syllable contact clusters by \citet[ chapter 3]{Hammond1999a} and \citet[ chapter 8]{Duanmu2009}, a list of simplex words was derived from the ``lemmas'' (that is, uninflected forms) in the English portion of CELEX \citep{CELEX}, a database of morphological, phonological, and syntactic annotations based on the COBUILD corpus, a resource that was also used to construct the Collins English dictionary employed by \citeauthor{Pierrehumbert1994}.

Two further filters were applied to this list. First, any lemma coded as an unassimilated loanword is excluded. Secondly, all lemmas that have a ``morphological status'' other than ``monomorpheme'' are excluded. This is crucial because CELEX makes no guarantee that lemmas are simplex; indeed, many lemmas are products of derivation (e.g., \emph{abusive}, from \emph{abuse}). These stringent criteria have a profound effect on the makeup of the data. For instance, nearly all exceptions to \textsc{Obstruent Voice Assimilation} (see \S\ref{ova} below) cited by \citet[74]{Hammond1999a} are excluded as unassimilated loanwords (e.g., \emph{vodka}, \emph{smorgasbord}) or as morphologically complex (e.g., \emph{jurisdiction}, \emph{madcap}, \emph{tadpole}, \emph{scapegoat}, \emph{magpie}).

``Neo-classical compounds'', words that appear to consist of a Latinate prefix and a bound stem, like \emph{inspect} or \emph{excrete}, have an ambiguous and long-debated morphological status. \citeauthor{Pierrehumbert1994} attempts to distinguish between words like \emph{complete}, \emph{extreme}, \emph{inspect}, \emph{obtuse}, which she assumes to be simplex, and \emph{excrete} and \emph{transparent}, which she considers complex. However, CELEX codes these words as complex and they are thus excluded. This is consistent with converging evidence that speakers decompose these Latinate forms. The baroque treatment of Latinate forms in \emph{The sound pattern of English} (\citealp{SPE}, henceforth \emph{SPE}) assumes a prefix/bound stem decomposition to simplify various morphophonological generalizations.  In a similar vein, \citet[11--13]{Aronoff1976} observes that Latinate forms which share the same bound stem also share irregular allomorphs of that stem under derivation, which is presented as evidence of prefix/bound stem decomposition.

%\footnote{Similar behaviors are found in Russian verbs which share roots but which take different prefixes or the reflexive \citep{Pesetsky1977}.}

\begin{example}[Bound stem-specific allomorphy]
\begin{tabular}{l l l l l l}
a. & {adhere}   & {adhesion}   \\
   & {cohere}   & {cohesion}   \\
b. & {conceive} & {conception} \\
   & {perceive} & {perception} \\
\end{tabular}
\end{example}

There is also a syntactic interaction between Latinate prefixes on verbs and the makeup of the verbal complement. \citet{Harley2009} observes that Latinate verbs do not generally participate in ditransitive, verb participle, or adjectival resultative constructions, all acceptable with similar Anglo-Saxon verbs \citep[see also][]{Gropen1989,Coppock2008}. \citeauthor{Harley2009} proposes that the Latinate prefix is introduced in a low small clause, the same position as the theme, participle, and result adjectival in the Anglo-Saxon examples.

\begin{example}[Latinate verbs and small clauses] \label{harley}
\begin{tabular}{ll@{}ll@{}l}
a. & * & {exhibit him the painting} & ~ & {show him the painting} \\
b. & * & {imbibe himself stupid}    & ~ & {drink himself stupid}  \\
c. & * & {exhibit it off}           & ~ & {show it off}           \\
\end{tabular}
\end{example}

Decomposition is also supported by lexical decision results. Among nonce words, those that can be exhaustively decomposed into prefix and bound stem, like *\emph{de}-\emph{juvenate}, are processed slower that contain a licit morph but cannot be exhaustively decomposed, like *\emph{de-pertoire}, and these in turn are processed slower than those which admit no such decomposition \citep{Taft1975}.\footnote{The same contrasts are present in many other domains where decomposition into root and affix(es) is less controversial, for instance, inflected verbs in Italian \citep{Caramazza1988}.} \citet{Emmorey1989} reports facilitative auditory priming between pairs of words which share a bound stem, like \emph{permit}-\emph{submit} \citep[though see][]{Marslen-Wilson1994}.\footnote{It may be the case that full decomposition is only one of the two methods for the processing of complex words, the other being whole word looking, either running in serial \citep{Caramazza1988} or in parallel \citep{Baayen1997b}. Whatever evidence that is amassed in support of a whole word lookup procedure, however, does not undermine the considerable evidence for decomposition.}

\subsubsection{Syllabification and phonologization}

CELEX includes broad, syllabified Received Pronunciation transcriptions of individual wordforms. However, a casual inspection of the data reveals the unsystematic nature of these syllabifications. Any two words sharing a nucleus-medial cluster-nucleus sequence should be given the same parse, since syllabification is universally non-contrastive. Yet many putative syllabification contrasts can be found in CELEX; for instance, compare the parses given to the clusters in \emph{chemistry} [ˈkɛ.mɪ.stɹɪ] and \emph{ministry} [ˈmɪ.nɪs.tɹɪ].\footnote{Note that word-final \emph{y} is generally lax [ɪ] in Received Pronunciation \citep[][II.294]{AOE}.}

A systematic syllabification requires an automated procedure for delimiting medial clusters and parsing them into coda and onset. Since this procedure is applied only to medial clusters, this study remains agnostic on several contentious issues concerning syllabification of English; for instance, there is no need to address the status of so-called ``ambisyllabic'' consonants, since this does not occur with medial clusters, or morphological effects on syllabification, since no complex words are included. The technique used here is a variation on the theme of onset maximization (e.g., \citealp[42f.]{Kahn1976}, \citealp{Kurylowicz1948}, \citealp[75]{Pulgram1970}, \citealp[][358f.]{Selkirk1982b}) which favors parses of word-medial clusters in which as much of the cluster as possible is assigned to the onset. Initial onsets are used to define what is a ``possible'' medial onset \citep[though cf.][36]{Fischer-Jorgensen1952}. Medial clusters in words like \emph{neu}[.tɹ]\emph{on} or \emph{bi}[.stɹ]\emph{o} are found in word-initial position (e.g., [tɹ]\emph{ansit}, [stɹ]\emph{ike}), so onset maximization assigns the entire medial cluster to the onset, leaving the medial coda empty. In contrast, the cluster in \emph{mi}[n.stɹ]\emph{el} is not found word-initially; the maximal onset here is [stɹ], and the residual [n] is assigned to the coda.

\paragraph{Stressed lax vowels} When a medial consonant cluster is preceded by a stressed lax vowel, as in words like \emph{wh}[ɪs.p]\emph{er}, \emph{v}[ɛs.t]\emph{ige}, or \emph{m}[ʌs.k]\emph{et}, the first consonant of the cluster checks the lax vowel \citep[e.g.,][3]{Hammond1997}. \citet[][55]{Harris1994} notes that onset maximization produces an incorrect result when, in addition, the medial cluster is a valid onset: in \emph{whisper}, \emph{vestige}, and \emph{musket}, onset maximization incorrectly assigns [sp, st, sk] to the medial onset, leaving the lax vowel unchecked. Consequently, the first consonant of a medial cluster is assigned to the coda of a preceding syllable before a stressed lax vowel \citep[cf.][48]{Pulgram1970}, a minimal modification of the onset maximization parse.

\paragraph{Affricates} If the affricates [tʃ, dʒ] are treated as stop-fricative sequences \citep[e.g.,][]{Hualde1988,Lombardi1990}, the previous principle will incorrectly assign the two segments of the affricate to separate syllables before stressed lax vowels (e.g., *\emph{ra}[t.ʃ]\emph{et} or \emph{a}[d.ʒ]\emph{ile}) unless some ad hoc constraint \citep[e.g.,][]{Wells1990} is in place. A more traditional analysis of affricates, in which they are simple segments distinguished from pure stops by stridency \citep[24]{Jakobson1961} or delayed release (\emph{SPE}:321f.) is assumed here, following \citeauthor{Pierrehumbert1994}. This assumption can be motivated by the tendency of affricates to pattern with individual segments in phonotactic generalizations. For instance, Classical Nahua allows the affricate series [ts, tʃ, tɬ] in onsets, but bans all other onset clusters \citep[9]{Launey2011}. 
%\citet[86]{Butskhrikidze2002} notes a similar generalization in the consonant clusters of Georgian.

\paragraph{The velar nasal} According to some \citep[e.g.,][]{Sapir1925}, [ŋ] in English is a phoneme in its own right, but others (e.g., \emph{SPE}:85, \citealp{Borowsky1986}:65f.) hold that it is simply the allophone of /n/ before underlying /k, ɡ/ (with later deletion of /ɡ/ in some contexts). The total absence of onset [ŋ], where it cannot be followed by a dorsal consonant which conditions the velar allophone, is strong evidence that [ŋ] is a pure allophone of /n/, the analysis assumed here.
